\textbf{ Interpretable Gaussian Processes for Stellar Light Curves }

### 2020-07-29 tex/ms.tex
more text

%
\section{The Spot Expansion}
\label{sec:spot}
%
We adopt the following expression for the spherical harmonic coefficient
of degree $l$ and order $m$ in the expansion of a spot at
$\theta = \varphi = 0$:
%
\begin{align}
    \label{eq:ylm0}
    y_{lm}(\delta, r, \theta = 0, \varphi = 0) =
    \begin{cases}
        1 - \dfrac{\delta c r}{2 (1 + c r)}
         & l = m = 0    \\[2em]
        -\dfrac{\delta c r \left( 2 + c r \right)}
        {2 \sqrt{2l + 1} (1 + c r)^{l + 1}}
         & l > 0, m = 0 \\[2em]
        0
         & m \ne 0
    \end{cases}
\end{align}
%
where $\delta \in [-\infty, 1]$ is the fractional decrease in the brightness
at the center of the spot and $r \in [0, 1]$ is a normalized spot radius.
The quantity $c$ is a normalization constant for the radius (see below).

The expression in Equation~(\ref{eq:ylm0}) is convenient because it satisfies
three important properties:
%
\begin{enumerate}[itemsep=2pt,parsep=1pt,label=\textbf{\arabic*}]
    \item The surface intensity monotonically increases away from the spot center
    \item The surface intensity at the spot center is $1 - \delta$
    \item The surface intensity at the antipode of the spot center is unity
\end{enumerate}
%
These properties may be demonstrated by considering the expression for the
surface intensity at a given point $(\theta, \varphi)$:
%
\begin{align}
    I(\theta, \varphi)
     & =
    \sum\limits_{l=0}^\infty \sum\limits_{m=-l}^l
    y_{lm} Y_{lm}(\theta, \varphi) \\
     & = \sum\limits_{l=0}^\infty
    y_{l0} \sqrt{2l + 1} P_l(\cos\theta)
\end{align}
%
where $P_l$ is the Legendre polynomial of degree $l$ and we have implicitly
assumed a normalization such that the integral of our expansion over
the unit sphere is $4\pi$.
Combining this with Equation~(\ref{eq:ylm0}) and rearranging, we may write
%
\begin{align}
    \label{eq:Igen}
    I(\theta, \varphi) =
    1 + \dfrac{\delta c r}{2}
    -
    \dfrac{\delta c r \left( 2 + c r \right)}{2 (1 + c r)}
    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + c r}\right)^l P_l(\cos\theta)
    \quad.
\end{align}
%
The summation in Equation~(\ref{eq:Igen}) has a closed-form expression in
terms of the generating function of the Legendre polynomials:
%
\begin{align}
    \label{eq:gen}
    \sum\limits_{l=0}^\infty t^l P_l(\cos\theta) = \frac{1}{\sqrt{1 + t^2 - 2 t \cos\theta}}
    \quad,
\end{align}
%
so we may express the intensity in the fairly simple form
%
\begin{align}
    \label{eq:Ifinal}
    I(\theta, \varphi) = A - \frac{B}{\sqrt{C - \cos\theta}}
    \quad,
\end{align}
%
where
%
\begin{align}
    A & = 1 + \dfrac{\delta c r}{2}                      \\
    B & = \delta c r (2 + cr) \sqrt{\dfrac{1}{8 + 8 cr}} \\
    C & = \dfrac{1 + (1 + c r)^2}{2 + 2 c r}
\end{align}
%
are positive constants.

Differentiating Equation~(\ref{eq:Ifinal}) with respect to $\theta$, we have
%
\begin{align}
    \label{eq:Ideriv}
    \dfrac{\mathrm{d}I(\theta, \varphi)}{\mathrm{d}\theta} & =
    -\frac{B\sin\theta}{2(C - \cos\theta)^\frac{3}{2}}
    \quad,
\end{align}
%
which is zero only for $\theta = 0$ (for which $I(\theta, \varphi)$ is
minimized) and $\theta = \pi$ (for which it is maximized). The intensity
therefore increases monotonically from the spot center to the antipode,
as stated in \textbf{1}. The value at the minimum is
%
\begin{align}
    I_{\mathrm{min}} & = A - \dfrac{B}{\sqrt{C - 1}} \\
                     & = 1 - \delta
    \quad,
\end{align}
%
as stated in \textbf{2}, and the value at the maximum is
%
\begin{align}
    I_{\mathrm{max}} & = A - \dfrac{B}{\sqrt{C + 1}} \\
                     & = 1
    \quad,
\end{align}
%
as stated in \textbf{3}.

% At the spot center, $\theta = 0$, so $\cos\theta = 1$. Since $P_l(1) = 1$ for
% all $l$, we have
% %
% \begin{align}
%     I(0)
%      & =
%     \sum\limits_{l=0}^\infty y_{l0} \sqrt{2l + 1} \\
%      & =
%     1 - \dfrac{\delta c r}{2 (1 + c r)}
%     -
%     \dfrac{\delta c r (2 + c r)}{2 (1 + c r)}
%     \sum\limits_{l=1}^\infty
%     \dfrac{1}
%     {(1 + c r)^{l}}                               \\
%      & =
%     1 -
%     \delta
%     \left[
%         \dfrac{c r}{2 (1 + c r)}
%         - \dfrac{2 + c r}{2(1 + c r)}
%         \right]                                   \\
%      & = 1 - \delta
%     \quad,
% \end{align}
% %
% as stated in \textbf{2}, where we made use of the fact that
% %
% \begin{align}
%     \sum\limits_{l=1}^\infty
%     \dfrac{1}
%     {(1 + x)^{l}} = \frac{1}{x}
%     \quad.
% \end{align}
% %
% Finally, at the antipode, $\theta = \pi$,
% $\cos\theta = -1$, and $P_l(\cos\theta) = (-1)^l$, so
% %
% \begin{align}
%     I(\pi)
%      & =
%     \sum\limits_{l=0}^\infty y_{l0} (-1)^l \sqrt{2l + 1} \\
%      & =
%     1 - \dfrac{\delta c r}{2 (1 + c r)}
%     -
%     \dfrac{\delta c r (2 + c r)}{2 (1 + c r)}
%     \sum\limits_{l=1}^\infty
%     \dfrac{(-1)^l}
%     {(1 + c r)^{l}}                                      \\
%      & =
%     1 - \dfrac{\delta c r}{2 (1 + c r)}
%     +
%     \dfrac{\delta c r}{2 (1 + c r)}                      \\
%      & = 1
%     \quad,
% \end{align}
% as stated in \textbf{3}, where we made use of the fact that
% %
% \begin{align}
%     \sum\limits_{l=1}^\infty
%     \dfrac{(-1)^l}
%     {(1 + x)^{l}} = -\frac{1}{2 + x}
%     \quad.
% \end{align}
%
%\bibliography

>>>>

%
\section{The Spot Expansion}
\label{sec:spot}
%
\subsection{Preliminaries}
%
We adopt the following expression for the spherical harmonic coefficient
of degree $l$ and order $m$ in the expansion of a spot at
$\theta = \varphi = 0$:
%
\begin{align}
    \label{eq:ylm0}
    y_{lm}(r, \delta) =
    \begin{cases}
        1 - \dfrac{\delta c r}{2 (1 + c r)}
         & l = m = 0    \\[2em]
        -\dfrac{\delta c r \left( 2 + c r \right)}
        {2 \sqrt{2l + 1} (1 + c r)^{l + 1}}
         & l > 0, m = 0 \\[2em]
        0
         & m \ne 0
    \end{cases}
\end{align}
%
where $\delta \in [-\infty, 1]$ is the spot contrast (the fractional
decrease in the brightness
at the center of the spot) and $r \in (0, 1]$ is the normalized spot radius.
The quantity $c \in (0, \infty)$ is a normalization constant for the radius
(more details below).
The expression in Equation~(\ref{eq:ylm0}) is convenient because it satisfies
four important properties:
%
\begin{enumerate}[itemsep=2pt,parsep=1pt,label=\textbf{\arabic*}]
    \item The surface intensity monotonically increases away from the spot center
    \item The surface intensity at the spot center is $1 - \delta$
    \item The surface intensity at the antipode of the spot center is unity
    \item The size of the spot increases monotonically with $r$
\end{enumerate}
%
These properties may be demonstrated by considering the expression for the
surface intensity at a given point $(\theta, \varphi)$:
%
\begin{align}
    \begin{split}
        I(\theta, \varphi)
        & =
        \sum\limits_{l=0}^\infty \sum\limits_{m=-l}^l
        y_{lm} Y_{lm}(\theta, \varphi) \\
        & = \sum\limits_{l=0}^\infty
        y_{l0} \sqrt{2l + 1} P_l(\cos\theta)
    \end{split}
\end{align}
%
where $P_l$ is the Legendre polynomial of degree $l$ and we have implicitly
assumed a normalization such that the integral of our expansion over
the unit sphere is $4\pi$.
Combining this with Equation~(\ref{eq:ylm0}) and rearranging, we may write
%
\begin{align}
    \label{eq:Igen}
    I(\theta, \varphi) =
    1 + \dfrac{\delta c r}{2}
    -
    \dfrac{\delta c r \left( 2 + c r \right)}{2 (1 + c r)}
    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + c r}\right)^l P_l(\cos\theta)
    \quad.
\end{align}
%
The summation in Equation~(\ref{eq:Igen}) has a closed-form expression in
terms of the generating function of the Legendre polynomials:
%
\begin{align}
    \label{eq:gen}
    \sum\limits_{l=0}^\infty t^l P_l(\cos\theta) = \frac{1}{\sqrt{1 + t^2 - 2 t \cos\theta}}
    \quad,
\end{align}
%
so we may express the intensity in the fairly simple form
%
\begin{align}
    \label{eq:Ifinal}
    I(\theta, \varphi) & = A - \frac{B}{\sqrt{C - \cos\theta}}
    \quad,
    %
    \\
    \intertext{where}
    %
    \begin{split}
        A & = 1 + \dfrac{\delta c r}{2}                      \\
        B & = \delta c r (2 + cr) \sqrt{\dfrac{1}{8 + 8 cr}} \\
        C & = \dfrac{1 + (1 + c r)^2}{2 + 2 c r}
    \end{split}
\end{align}
%
are positive constants.

Differentiating Equation~(\ref{eq:Ifinal}) with respect to $\theta$, we have
%
\begin{align}
    \label{eq:Ideriv}
    \dfrac{\mathrm{d}I(\theta, \varphi)}{\mathrm{d}\theta} & =
    -\frac{B\sin\theta}{2(C - \cos\theta)^\frac{3}{2}}
    \quad,
\end{align}
%
which is zero only for $\theta = 0$ (for which $I(\theta, \varphi)$ is
minimized) and $\theta = \pi$ (for which it is maximized). The intensity
therefore increases monotonically from the spot center to the antipode,
as stated in \textbf{1}. The value at the minimum is
%
\begin{align}
    \begin{split}
        I_{\mathrm{min}} & = A - \dfrac{B}{\sqrt{C - 1}} \\
        & = 1 - \delta
        \quad,
    \end{split}
\end{align}
%
as stated in \textbf{2}, and the value at the maximum is
%
\begin{align}
    \begin{split}
        I_{\mathrm{max}} & = A - \dfrac{B}{\sqrt{C + 1}} \\
        & = 1
        \quad,
    \end{split}
\end{align}
%
as stated in \textbf{3}.
Finally, to show \textbf{4}, let us compute the half width at half minimum
$\Delta\theta$ of the intensity profile:
%
\begin{align}
    1 - \left(A - \dfrac{B}{\sqrt{C - \cos{\Delta\theta}}}\right) =
    \dfrac{1}{2}\delta
\end{align}
%
Solving for $\Delta\theta$ yields
%
\begin{align}
    \Delta\theta =
    \cos^{-1} \left[ \dfrac{2 + 3 c r (2 + c r)}{2 (1 + c r)^3} \right]
    \quad.
\end{align}
%
Differentiation with respect to $r$ yields
%
\begin{align}
    \dfrac{\mathrm{d}\Delta\theta}{\mathrm{d}r} =
    \frac{3 c^2 r \left(2 + c r\right)}{\left(1 + c r\right)^4
        \sqrt{\frac{c^2 r^2 \left(2 + c r\right)^2 \left(1 + 2 c r\right)
                \left(3 + 2 c r\right)}{\left(1 + c r\right)^6}}}
    \quad,
\end{align}
%
which is positive definite for all $c, r > 0$.

\subsection{Distributions over $r$ and $\delta$}
%
We model the normalized spot radius $r$ as a random variable drawn from
a Beta distribution, whose PDF is given by
%
\begin{align}
    p \big(r \, \big| \, \alpha, \beta \big)
     & =
    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
    r^{\alpha - 1}
    (1 - r)^{\beta - 1}
    \quad,
\end{align}
%
where $\Gamma$ is the Gamma function and $\alpha$ and $\beta$ are
hyperparameters characterizing the shape of the distribution.
%
Since we require the brghtness $b = 1 - \delta$ at the center of the spot to
be non-negative, it is convenient to parametrize the distribution of $b$
instead of $\delta$. We model $b$ as a random variable drawn from a log-normal
distribution:
%
\begin{align}
    p \big(b \, \big| \, \mu, \nu \big)
     & =
    \frac{1}{b\sqrt{2\pi\nu}}
    \exp\left[
        -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
        \right]
    \quad,
\end{align}
%
where $\mu$ and $\nu$ are the mean and variance of the log-normal, respectively.
The joint distribution over $r$ and $b$ is therefore separable,
by construction:
%
\begin{align}
    p \big( r, b \, \big| \, \alpha, \beta, \mu, \nu \big)
     & =
    p \big(r \, \big| \, \alpha, \beta \big)
    \,
    p \big(b \, \big| \, \mu, \nu \big)
    \quad.
\end{align}
%

\subsection{First moment}
%
The first moment of the distribution of $y_{lm}$ over $r$ and $b$ is
%
\begin{align}
    \label{eq:Erb}
    \begin{split}
        \mathrm{E} \big[ y_{lm} \big] &=
        \int\limits_0^\infty \int\limits_0^1
        y_{lm} (r, 1 - b)
        p \big(r, 1 - b \, \big| \, \alpha, \beta, \mu, \nu \big)
        \mathrm{d} r
        \,
        \mathrm{d} b
        \\
        &=
        \int\limits_0^\infty
        \mathrm{E}_r \big[ y_{lm} \big]
        p \big(b \, \big| \, \mu, \nu \big)
        \mathrm{d} b
        \quad,
    \end{split}
    %
    \\
    \intertext{where}
    \label{eq:Er}
    %
    \mathrm{E}_r \big[ y_{lm} \big]
     & =
    \int\limits_0^1
    y_{lm} (r, 1 - b)
    p \big(r \, \big| \, \alpha, \beta \big)
    \mathrm{d} r
    \quad.
\end{align}
%
Expanding Equation~(\ref{eq:Er}), we obtain
%
\begin{align}
    \mathrm{E}_r \big[ y_{lm} \big]
     & =
    \begin{dcases}
        \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
        \int\limits_0^1
        \left(
        1 - \dfrac{(1 - b) c r}{2 (1 + c r)}
        \right)
        r^{\alpha - 1}
        (1 - r)^{\beta - 1}
        \mathrm{d} r
         &
        \qquad
        l = m = 0    \\[2em]
        -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
        \int\limits_0^1
        \dfrac{(1 - b) c r \left( 2 + c r \right)r^{\alpha - 1}
        (1 - r)^{\beta - 1}}
        {2 \sqrt{2l + 1} (1 + c r)^{l + 1}}
        \mathrm{d} r
         &
        \qquad
        l > 0, m = 0 \\[2em]
        0
         &
        \qquad m \ne 0
        \quad.
    \end{dcases}
\end{align}
%
Thanks to the dependence of the spherical harmonic coefficients on only
powers of $r$ and $(1 + cr)$, the integrals above may be expressed in closed
form in terms of the hypergeometric function ${_2F_1}$:
%
\begin{align}
    \mathrm{E}_r \big[ y_{lm} \big]
     & =
    \begin{dcases}
        1 -
        \dfrac{(1 - b) \alpha c}{2(\alpha + \beta)} F_l
         &
        \qquad
        l = m = 0    \\[2em]
        -\dfrac{(1 - b) \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
        \left[
            F_l + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l
            \right]
         &
        \qquad
        l > 0, m = 0 \\[2em]
        0
         &
        \qquad m \ne 0
    \end{dcases}
\end{align}
%
where we define
%
\begin{align}
    \begin{split}
        F_l & \equiv {_2F_1}(l + 1, \alpha + 1; \alpha + \beta + 1; -c)
        \\[1em]
        G_l & \equiv {_2F_1}(l + 1, \alpha + 2; \alpha + \beta + 2; -c)
        \quad.
    \end{split}
\end{align}
%
Note that for a given value of $\alpha$, $\beta$, and $c$, we need only compute
$F_0$, $F_1$, $G_0$, and $G_1$ directly, since the remaining terms may be
obtained recursively:
%
\begin{align}
    \begin{split}
        F_l &=
        \dfrac{\alpha + \beta + 1 - l}{2l}F_{l - 2}
        - \dfrac{2\alpha + \beta + 2 - 3l}{2l}F_{l - 1}
        \\[1em]
        G_l &=
        \dfrac{\alpha + \beta + 2 - l}{2l}G_{l - 2}
        - \dfrac{2\alpha + \beta + 4 - 3l}{2l}G_{l - 1}
        \quad.
    \end{split}
\end{align}
%
We may now evaluate the integral in Equation~(\ref{eq:Erb}):
%
\begin{align}
    \mathrm{E} \big[ y_{lm} \big] & =
    \begin{dcases}
        \frac{1}{b\sqrt{2\pi\nu}}
        \int\limits_0^\infty
        \left(
        1 -
        \dfrac{(1 - b) \alpha c}{2(\alpha + \beta)} F_l
        \right)
        \\
        \qquad\qquad\times
        \exp\left[
            -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
            \right]
        \mathrm{d} b
         &
        \qquad\qquad
        l = m = 0
        \\
        -
        \frac{1}{b\sqrt{2\pi\nu}}
        \int\limits_0^\infty
        \dfrac{(1 - b) \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
        \\
        \qquad\qquad\times
        \left[
            F_l + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l
            \right]
        \\
        \qquad\qquad\times
        \exp\left[
            -\dfrac{\left(\ln b - \mu\right)^2}{2\nu}
            \right]
        \mathrm{d} b
         &
        \qquad\qquad
        l > 0, m = 0 \\[2em]
        0
         &
        \qquad\qquad m \ne 0
    \end{dcases}
\end{align}
%
Most of the terms in the integrands above are constants, so these
integrals may also be expressed in closed form:
%
\begin{align}
    \mathrm{E} \big[ y_{lm} \big]
     & =
    \begin{dcases}
        1 -
        \dfrac{\gamma \alpha c}{2(\alpha + \beta)} F_l
         &
        \qquad
        l = m = 0    \\[2em]
        -\dfrac{\gamma \alpha c}{\sqrt{2l + 1} (\alpha + \beta)}
        \left[
            F_l + \dfrac{(\alpha + 1)c}{2(\alpha + \beta + 1)} G_l
            \right]
         &
        \qquad
        l > 0, m = 0 \\[2em]
        0
         &
        \qquad m \ne 0
    \end{dcases}
\end{align}
%
where we define
%
\begin{align}
    \gamma \equiv 1 - \exp\left[ \mu + \frac{1}{2}\nu\right]
    \quad.
\end{align}
%

\subsection{Second moment}
%
The second moment of the distribution of $y_{lm}$ over $r$ and $b$ is
%
\begin{align}
    \label{eq:Erb2}
    \begin{split}
        \mathrm{E} \big[ y_{lm} y_{l'm'} \big] &=
        \int\limits_0^\infty \int\limits_0^1
        y_{lm} (r, 1 - b)
        y_{l'm'} (r, 1 - b)
        p \big(r, 1 - b \, \big| \, \alpha, \beta, \mu, \nu \big)
        \mathrm{d} r
        \,
        \mathrm{d} b
        \\
        &=
        \int\limits_0^\infty
        \mathrm{E}_r \big[ y_{lm} y_{l'm'} \big]
        p \big(b \, \big| \, \mu, \nu \big)
        \mathrm{d} b
        \quad,
    \end{split}
    %
    \\
    \intertext{where}
    \label{eq:Er2}
    %
    \mathrm{E}_r \big[ y_{lm} y_{l'm'} \big]
     & =
    \int\limits_0^1
    y_{lm} (r, 1 - b)
    y_{l'm'} (r, 1 - b)
    p \big(r \, \big| \, \alpha, \beta \big)
    \mathrm{d} r
    \quad.
\end{align}
%
Expanding Equation~(\ref{eq:Er2}), we obtain
%
\begin{align}
    \mathrm{E}_r \big[ y_{lm} y_{l'm'} \big]
     & =
    \begin{dcases}
        \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
        \int\limits_0^1
        \left(
        1 - \dfrac{(1 - b) c r}{2 (1 + c r)}
        \right)^2
        r^{\alpha - 1}
        (1 - r)^{\beta - 1}
        \mathrm{d} r
         &
        \quad\quad
        \parbox[t]{\linewidth}{$l = l',              \\m = m' = 0$}
        \\[2em]
        %
        -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
        \int\limits_0^1
        \left(1 - \dfrac{(1 - b) c r}{2 (1 + c r)}\right)
        \\
        \qquad\qquad\quad\quad\times
        \dfrac{(1 - b) c r \left( 2 + c r \right) r^{\alpha - 1}
            (1 - r)^{\beta - 1}}
        {2 \sqrt{2l + 1} (1 + c r)^{l + 1}}
        \mathrm{d} r
         &
        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-3em}$l > 0,  \\l' = 0, \\m = m' = 0$}
        \\[2em]
        %
        -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
        \int\limits_0^1
        \left(1 - \dfrac{(1 - b) c r}{2 (1 + c r)}\right)
        \\
        \qquad\qquad\quad\quad\times
        \dfrac{(1 - b) c r \left( 2 + c r \right) r^{\alpha - 1}
            (1 - r)^{\beta - 1}}
        {2 \sqrt{2l' + 1} (1 + c r)^{l' + 1}}
        \mathrm{d} r
         &
        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-3em}$l = 0,  \\l' > 0, \\m = m' = 0$}
        \\[2em]
        %
        -\dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
        \int\limits_0^1
        \dfrac{(1 - b)^2 c^2 r^2 \left( 2 + c r \right)^2 r^{\alpha - 1}
            (1 - r)^{\beta - 1}}
        {4 \sqrt{(2l + 1)(2l' + 1)} (1 + c r)^{l + l' + 2}}
        \mathrm{d} r
         &
        \quad\quad
        \parbox[t]{\linewidth}{\vspace{-1.5em}$l > 0 \\l' > 0 \\m = m' = 0$}
        \\[2em]
        %
        0
         &
        \quad\quad m, m' \ne 0
        \quad.
    \end{dcases}
\end{align}

%\bibliography
